---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis-CI Build Status](https://travis-ci.org/cboettig/template.svg?branch=master)](https://travis-ci.org/cboettig/template) [![Coverage Status](https://coveralls.io/repos/cboettig/template/badge.svg)](https://coveralls.io/r/cboettig/template)



This repository provides the current template I use for new research projects.  

## Why an R package structure?

Academic research isn't software development, and there are many other templates for how to organize a research project.  So why follow an R package layout?  Simply put, this is because the layout of an R package is familiar to a larger audience and allows me to leverage a rich array of tools that don't exist for more custom approaches.

"But"", you say, "a paper doesn't have unit tests, or documented functions!  Surely that's a lot of needless overhead in doing this!?"

Exactly... 

While there is certainly no need to use all the elements of a package in every research project, or even to have a package that can pass `devtools::check()` or even `devtools::install()` for it to be useful.  Most generic layout advice starts sounding like an R package pretty quickly: have a directory for `data/`, a separate one for `R/` scripts, another one for the manuscript files, and so forth.  [Temple Lang and Gentleman (2007)] advance the proposal for using the R package structure as a "Research Compendium," an idea that has since caught on with many others.

As a project grows in size and collaboration, having the more rigid structure offered by a package format becomes increasingly valuable. Packages can automate installation of dependencies, perform checks that changes have not broken code, and provide a modular and highly tested framework for collecting together the three essential elements: data, code, and manuscript-quality documentation, in a powerful and feature-rich environment.


[Temple Lang and Gentleman (2007)]: "http://doi.org/10.1198/106186007X178663"



## Steps to create the template

To use this template, I will usually clone this repo and just remove the `.git` record, starting off a new project accordingly.  Here I document the steps used to set up this template from scratch, which permits a slightly more modular approach.  If this were fully automated it would be preferable to copying, but has not yet reached that stage.


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

The template can be initialized with functions from devtools:

```{r, eval=FALSE}
devtools::install_github("hadley/devtools")
library("devtools")
```

Configure some default options for `devtools`, see `package?devtools`:

```{r, eval=FALSE}
options(devtools.name = "Carl Boettiger", 
        devtools.desc.author = "person('Carl', 'Boettiger', email='cboettig@gmail.com', role = c('aut', 'cre'))",
        devtools.desc.license = "MIT + file LICENSE")
```


Run devtools templating tools

```{r, eval=FALSE}
setup()
use_testthat()
use_vignette("intro")
use_travis()
use_package_doc()
use_cran_comments()
use_readme_rmd()
```

Additional modifications and things not yet automated by `devtools`: 

- Add the now-required LICENSE template data
- add `covr` to the suggests list


```{r, eval=FALSE}
writeLines(paste("YEAR: ", format(Sys.Date(), "%Y"), "\n", 
                 "COPYRIGHT HOLDER: ", getOption("devtools.name"), sep=""),
           con="LICENSE")

use_package("covr", "suggests")

write(
"
r_binary_packages:
  - testthat
  - knitr

r_github_packages:
  - jimhester/covr

after_success:
  - Rscript -e 'library(covr); coveralls()'",
file=".travis.yml", append=TRUE)
```

### Further steps that aren't automated

Further steps aren't yet automated in devtools or by me; as it's easier to add these manually to the template and then use the template when starting a new project.  


- add the travis shield to README, (as prompted to do by `add_travis()`)
- Turn on repo at coveralls.io and add the shield to README
- adding additional dependencies to DESCRIPTION with `use_package`, and also add to `.travis.yml` manually, e.g. under `r_binary_packages:`, `r_github_packages`, or `r_packages`
- add additional data with `use_data()` or possibly `use_raw_data()` (for scripts that import and clean data first)

## Manuscript elements

- Recent developments in `rmarkdown`, `knitr` and `rticles` packages greatly faciliates using vignettes as full manuscripts.  The above step adds only a basic HTML templated vignette.  This package includes a template for a latex/pdf manuscript using these tools. The actual template appropriate for a project may be better selected from (possibly my fork of) the `rticles` templates.

